---
title: "Maternal PFAS transfer in wild birds: a systematic review and meta-analysis"
author: "L.Ricolfi"
date: '2022-06-28'
output: html_document
editor_options: 
  chunk_output_type: console
---

# Project title

*Maternal PFAS transfer in wild birds: a systematic review and meta-analysis*

# Purpose

This .Rmd document provides the code we used for data processing and visualization.

# Data model

We use a dm object to create a relational data model from our local data tables. We extracted data from the included studies on six data tables:

-   **study_info**

-   **species_info**

-   **pfas_info**

-   **cohort_info**

-   **sample_info**

-   **measurements**

They are stored in [a repository on GitHub](https://github.com/ThisIsLorenzo/PFAS_MA_MaternalTransfer_WildBirds_Pilot/tree/main/Data)

# Data processing

```{r setup, eval = TRUE, include = TRUE, message = FALSE, echo = TRUE,  warning = FALSE}
# knitr setting
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  tidy = TRUE,
  cache = TRUE, 
  echo=TRUE
)

rm(list = ls())
```

```{r setup, results = 'hide'}
# Loading packages
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,
               readr,
               ape,
               curl,
               rotl,
               here,
               dm,
               dplyr,
               metafor,
               meta,
               orchaRd,
               kableExtra,
               clubSandwich)
```

```{r loading data, results = 'hide'}
st <- read_csv(here("Data", "study_info.csv"))
sp <- read_csv(here("Data", "species_info.csv"))
pfas <- read_csv(here("Data", "pfas_info.csv"))
co <- read_csv(here("Data", "cohort_info.csv"))
sa <- read_csv(here("Data", "sample_info.csv"))
me <- read_csv(here("Data", "measurement_info.csv"))
```

```{r cleaning data tables, results = 'hide'}
st <- st %>% 
  select(-("Timestamp")) %>%
  select(-(ends_with("comment")))
sp <- sp %>% 
  select(-("Timestamp")) %>%
  select(-(ends_with("comment")))
pfas <- pfas %>% 
  select(-("Timestamp")) %>% 
  select(-(ends_with("comment")))
co <- co %>% 
  select(-("Timestamp")) %>%
  select(-(ends_with("comment")))
sa <- sa %>% 
  select(-("Timestamp")) %>% 
  select(-(ends_with("comment")))
me <- me %>%
  select(-("Timestamp")) %>% 
  select(-(ends_with("comment")))
```

```{r adding tables to the dm object, results = 'hide'}
data_dm_no_keys <- dm(st, sp, pfas, co, sa, me)
data_dm_no_keys
data_dm_no_keys$st
data_dm_no_keys[c("st", "co")]
```

```{r defining primary keys, results = 'hide'}
data_dm_only_pks <- data_dm_no_keys %>% 
  dm_add_pk(table = st, columns = study_ID) %>% 
  dm_add_pk(sp, species_ID) %>% 
  dm_add_pk(pfas, pfas_ID) %>% 
  dm_add_pk(co, cohort_ID) %>% 
  dm_add_pk(sa, sample_ID) %>% 
  dm_add_pk(me, measurement_ID)
data_dm_only_pks
```

```{r defining foreign keys, results = 'hide'}
data_dm_all_keys <- 
  data_dm_only_pks %>% 
  dm_add_fk(table = co, columns = study_ID, ref_table = st) %>% 
  dm_add_fk(table = co, columns = species_ID, ref_table = sp) %>%
  dm_add_fk(table = sa, columns = cohort_ID, ref_table = co) %>%
  dm_add_fk(table = me, columns = sample_ID, ref_table = sa) %>%
  dm_add_fk(table = me, columns = pfas_ID, ref_table = pfas)
data_dm_all_keys
```

```{r visualization, results = 'hide'}
data_dm_all_keys %>% 
  dm_draw()
```

```{r integrity check, results = 'hide'}
data_dm_all_keys %>% 
  dm_examine_constraints()
```

The following chunk creates a new table that merges sa, co, and me. me includes only the variables we need to calculate effect sizes (i.e., min, max, median, etc. are removed)

```{r moving columns among dm objects}
me1 <- 
  data_dm_all_keys %>% 
  dm_select(me, measurement_ID | group_info | sample_ID | pfas_ID | mean_arithmetic | SD | n | limit_type) %>% 
  dm_select(sa, sample_ID : sample_type) %>% 
  dm_select(co, cohort_ID : data_pooled) %>% 
  dm_select(pfas, pfas_ID) %>% 
  dm_select(sp, species_ID) %>% 
  dm_select(st, study_ID) %>% 
  dm_flatten_to_tbl(.start = me, .recursive = TRUE, .join = left_join)

colnames(me1)
```

Now I want to create a new column called "es_ID" which matches all the pairs of measurement I will use in the escalc() function. Study, PFAS type, species, and sampling location must be the same, so I group by them. The group (i.e., group_info) must be different

```{r}
#group by

list_dat <- 
  me1 %>% 
  dplyr::group_split(study_ID,
                  pfas_ID , 
                  species_ID,
                  sampling_location)


# TODO 

# tdat <- data.frame(me1[me1$study_ID == "S_003" & me1$pfas_ID == "pfas_001", ])
# 
# 
# adult <- tdat[tdat$group_info == "adult", 1]
# progeny <-tdat[tdat$group_info == "progeny", 1]
# 
# ids <- expand.grid(adult_id = adult, progeny_id = progeny)
# 
# adult_id <- ids[[1]]
# progeny_id <- ids[[2]]
# 
# pos_adult <- match(adult_id, tdat$measurement_ID)
# adult_dat <- tdat[pos_adult, ]
# 
# pos_progeny <- match(progeny_id, tdat$measurement_ID)
# progeny_dat <- tdat[pos_progeny, ]
# 
# #ndat<-as.data.frame(cbind(adult_dat, chick_dat))
# 
# ndat <- data.frame(measurement_ID_A = adult_dat[["measurement_ID"]],
#            measurement_ID_C = progeny_dat[[ "measurement_ID"]],
#            mean_arithmetic_A = adult_dat[["mean_arithmetic"]],
#            mean_arithmetic_C =  progeny_dat[[ "mean_arithmetic"]]
#              )

# head(ndat)

# function

wide_table <-
  function(dat)
    {

dat <-
  as.data.frame(dat)  
  
adult <-
  dat[dat$group_info == "adult", 1]
progeny <-
  dat[dat$group_info == "progeny", 1]

ids <-
  expand.grid(adult_id = adult,
              progeny_id = progeny)

adult_id <- ids[[1]]
progeny_id <- ids[[2]]

pos_adult <- match(adult_id, dat$measurement_ID)
adult_dat <- dat[pos_adult, ]

pos_progeny <- match(progeny_id, dat$measurement_ID)
progeny_dat <- dat[pos_progeny, ]

#ndat<-as.data.frame(cbind(adult_dat, chick_dat))

ndat <- data.frame(
           measurement_ID_A = adult_dat[["measurement_ID"]],
           measurement_ID_P = progeny_dat[[ "measurement_ID"]],
           mean_arithmetic_A = adult_dat[["mean_arithmetic"]],
           mean_arithmetic_P =  progeny_dat[[ "mean_arithmetic"]],
           SD_A = adult_dat[["SD"]],
           SD_P = progeny_dat[["SD"]],
           n_A = adult_dat[["n"]],
           n_P = progeny_dat[["n"]],
           life_stage_A = adult_dat[["life_stage"]],
           life_stage_P = progeny_dat[["life_stage"]],
           study_ID_A = adult_dat[["study_ID"]],
           study_ID_P = progeny_dat[["study_ID"]],
           species_ID_A = adult_dat[["species_ID"]],
           species_ID_P = progeny_dat[["species_ID"]],
           sample_type_A = adult_dat[["sample_type"]],
           sample_type_P = progeny_dat[["sample_type"]],
           sampling_location_A = adult_dat[["sampling_location"]],
           sampling_location_P = progeny_dat[["sampling_location"]],
           pfas_ID_A = adult_dat[["pfas_ID"]],
           pfas_ID_P = progeny_dat[["pfas_ID"]],
           limit_type_A = adult_dat[["limit_type"]],
           limit_type_P = progeny_dat[["limit_type"]],
           sampling_year_A = adult_dat[["sampling_year"]],
           sampling_year_P = progeny_dat[["sampling_year"]],
           sampling_month_A = adult_dat[["sampling_month"]],
           sampling_month_P = progeny_dat[["sampling_month"]],
           developmental_stage_A = adult_dat[["developmental_stage"]],
           developmental_stage_P = progeny_dat[["developmental_stage"]],
           breeding_stage = adult_dat[["breeding_stage"]],
           laying_order = progeny_dat[["laying_order"]],
           nest = adult_dat[["nest"]],
           body_weight_mean = adult_dat[["body_weight_mean"]],
           body_weight_sd = adult_dat[["body_weight_sd"]],
           egg_weight_mean = progeny_dat[["egg_weight_mean"]],
           egg_weight_sd = progeny_dat[["egg_weight_sd"]],
           data_pooled = adult_dat[["data_pooled"]]
             )
ndat

}


dat <- list_dat[[3]]

res <- wide_table(list_dat[[3]])
res

# this applies wide_table data to every item in this list
ntable <- map_dfr(list_dat, wide_table)
```

Create effect sizes column and tidy data

```{r}
ntable <- dplyr::mutate(ntable, es_ID = as.vector(001 : 412)) # add esID column
ntable$es_ID <-  sub("(.{1})", "es_\\1", ntable$es_ID) #add es_IDs
# ntable <- ntable[, c(27, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26)] # move es_ID column at the beginning
# the following columns are created in order to link the new ntable to the dm abject
ntable <- dplyr::mutate(ntable, study_ID =  ntable$study_ID_A) # add study_ID column
ntable <- dplyr::mutate(ntable, species_ID =  ntable$species_ID_A) # add species_ID column
ntable <- dplyr::mutate(ntable, pfas_ID =  ntable$pfas_ID_A) # add pfas_ID column

ntable <- dplyr::select(ntable, -c('study_ID_A', 'study_ID_P', 'species_ID_A', 'species_ID_P', 'pfas_ID_A', 'pfas_ID_P'))
```

Remove effect sizes where both adult and progeny concentrations were under the limit of detection or quantification. This was an assumption made in the pre-registration.

```{r}
# Remove es_IDs where both limit_type_A AND limit_type_P are != NA
ntable <- 
  ntable %>% 
  dplyr::filter(is.na(limit_type_A) == "TRUE" |
               is.na(limit_type_P) == "TRUE")
```

Remove effect size measurements with NA in the mean concentration in the adult OR progeny. This is necessary for successful analysis because means cannot be NA.

```{r}
ntable <- 
  ntable %>% 
  dplyr::filter(is.na(mean_arithmetic_A) == "FALSE" & is.na(mean_arithmetic_P) == "FALSE")
```

#DM object

```{r adding the ntable to the dm object, include=FALSE}
nt <- ntable
data_dm_no_keys <- dm(st, sp, pfas, nt)

data_dm_only_pks <- data_dm_no_keys %>%
  dm_add_pk(table = st, columns = study_ID) %>%
  dm_add_pk(sp, species_ID) %>%
  dm_add_pk(pfas, pfas_ID) %>%
  dm_add_pk(nt, es_ID)
data_dm_only_pks

data_dm_all_keys <-
  data_dm_only_pks %>%
  dm_add_fk(table = nt, columns = study_ID, ref_table = st) %>% 
  dm_add_fk(table = nt, columns = species_ID, ref_table = sp) %>% 
  dm_add_fk(table = nt, columns = pfas_ID, ref_table = pfas)
data_dm_all_keys

data_dm_all_keys %>% 
  dm_draw()

data_dm_all_keys %>% 
  dm_examine_constraints()
```

Mergin the dm tables in a wide table

```{r merging all tables into one big table}
nt <-
  data_dm_all_keys %>%
  dm_flatten_to_tbl(.start = nt, .recursive = TRUE)
nt
```

# Functions

The `lnRR_func` function is here used to calculate a log response ratio (lnRR) adjusted for small sample sizes. In addition, this formula accounts for correlated samples. For more details, see *Doncaster and Spake (2018) Correction for bias in meta-analysis of little-replicated studies. Methods in Ecology and Evolution; 9:634-644*

```{r function to calculate effect size}
#Changing columns names to make them shorter
colnames(nt)[3] <- "mean_A"
colnames(nt)[4] <- "mean_P"
colnames(nt)[5] <- "sd_A"
colnames(nt)[6] <- "sd_P"

# Custom function to calculate the lnRR 
lnRR_func <- function(mean_A, n_A, mean_P, n_P, aCV2a, aCV2p){
  lnRR <- log(mean_P/mean_A) + 
        0.5 * ((aCV2p/n_P) - (aCV2a/n_A))	
  lnRR
}
# Custom function to calculate the lnRR's sampling variance from independent designs (rTC = 0)
var_lnRR_ind <- function(mean_A, n_A, mean_P, n_P, aCV2a, aCV2p){
  
  var_lnRR <- (aCV2a/n_A) + (aCV2p/n_P) 
  
  var_lnRR
}
# Mc: Concentration of PFAS of the raw (control) sample
# Nc: Sample size of the raw (control) sample
# Me: Concentration of PFAS of the cooked (experimental) sample
# Ne: Sample size of the cooked (experimental) sample 
# aCV2c: Mean coefficient of variation of the raw (control) samples
# aCV2e: Mean coefficient of variation of the cooked (experimental) samples

```

# Phylogenetic information

Import phylogenetic information and calculate phylogenetic variance-covariance matrix

NOTE: The phylogenetic tree was generated in the `tree_MA_MT_PFAS.Rmd` document

```{r, fig.height=10, fig.width = 8}
tree <- read.tree(here("R", "phylogenetic_tree.tre")) # Import phylogenetic tree (see tree_MA_MT_PFAS.Rmd for more details) 
tree <- compute.brlen(tree) # Generate branch lengths 
cor_tree <- vcv(tree,corr = T) # Generate phylogenetic variance-covariance matrix 

nt2 <- 
  nt

nt2$Phylogeny <- as.factor(str_replace(nt2$species_scientific_name, " ", "_")) # Add the `phylogeny` column to the data frame
colnames(cor_tree) %in% nt2$Phylogeny # Check correspondence between tip names and data frame

# Rename species names that do not match
levels(nt2$Phylogeny)[levels(nt2$Phylogeny) == "Larus_audouinii"] <- "Ichthyaetus_audouinii"
levels(nt2$Phylogeny)[levels(nt2$Phylogeny) == "Phalacrocorax_aristotelis"] <- "Gulosus_aristotelis"
levels(nt2$Phylogeny)[levels(nt2$Phylogeny) == "Diomedea_immutabilis"] <- "Phoebastria_immutabilis"

# checking all species are in the data
match(unique(nt2$Phylogeny),colnames(cor_tree))
match(nt2$Phylogeny, colnames(cor_tree))
# plotting tree
plot(tree)
```

# Effect sizes

Calculate effect sizes The average coefficient of variation in PFAS concentration was calculated for each study and treatment, according to *Doncaster and Spake (2018) Correction for bias in meta-analysis of little-replicated studies. Methods in Ecology and Evolution; 9:634-644*. Then, these values were averaged across studies and used to calculate the lnRR corrected for small sample sizes (for formula, see the `lnRR_func` above)

```{r Calculating lnRR and Variance}
# Calculate the squared coefficient of variation for adult and progeny groups

# colnames(nt2)[3] <- "mean_A"
# colnames(nt2)[4] <- "mean_P"
# colnames(nt2)[5] <- "sd_A"
# colnames(nt2)[6] <- "sd_P"

aCV2 <- nt2 %>% 
  group_by(study_ID) %>%
  summarise(CV2a = mean((sd_A/mean_A)^2, na.rm = T),
            CV2p = mean((sd_P/mean_P)^2, na.rm = T)) %>% 
  ungroup() %>% # ungroup 
  summarise(aCV2a = mean(CV2a, na.rm = T), # Mean CV^2 for exp and control groups across studies 
            aCV2p = mean(CV2p, na.rm = T))

lnRR <- lnRR_func(mean_A = nt2$mean_A,
                  n_A = nt2$n_A,
                  mean_P = nt2$mean_P,
                  n_P = nt2$n_P,
                  aCV2a = aCV2[[1]],
                  aCV2p = aCV2[[2]])

var_lnRR <- var_lnRR_ind(mean_A = nt2$mean_A,
                  n_A = nt2$n_A,
                  mean_P = nt2$mean_P,
                  n_P = nt2$n_P,
                  aCV2a = aCV2[[1]],
                  aCV2p = aCV2[[2]]) 

dat <- nt2 %>% 
  dplyr::mutate(N_tilde = (n_A*n_P)/(n_A + n_P)) # getting effective sample size

dat <- cbind(dat, lnRR, var_lnRR) # Merge effect sizes with the data frame
```

# Sensitivity analysis - Data

We'll create three data sets to run a sensitivity analysis. The first data set (i.e., dat) will include all the data. The second (i.e., dat2) one will drop out data that are from pooled samples (i.e., from adult male + female pooled data) and/or data from adults and progeny that come from different nests. The third one (i.e., dat3) will drop out all effect size measures where the PFAS concentration in the adult OR progeny is below the limit of detection/quantification.

```{r, eval=FALSE}
dat2 <-
  dat %>%
  dplyr::filter(data_pooled == "No" & nest == "Yes")
```

Remove all effect sizes where concentration in the adult OR progeny is below the limit of detection/quantification

```{r, eval=FALSE}
Remove es_IDs where limit_type_A OR limit_type_P is != NA
dat3 <-
  dat2 %>%
  dplyr::filter(if_all(c(limit_type_A, limit_type_P), ~is.na(.)))
```

# Distribution of effect sizes

```{r, fig.height=8, fig.width=15, , eval=FALSE}
mean
ggplot(dat, aes(x=lnRR))+ geom_histogram(fill = "salmon", col = "black", binwidth = 0.2) + theme_classic()
# variance
ggplot(dat, aes(x=var_lnRR))+ geom_histogram(fill = "salmon", col = "black", binwidth = 0.05) + theme_classic()
# log variance
ggplot(dat, aes(x=var_lnRR))+ geom_histogram(fill = "salmon", col = "black", binwidth = 0.05) + scale_x_log10()+theme_classic()
```

# Forest plot
```{r create the forest plot, , eval=FALSE}
forest(dat$lnRR, dat$var_lnRR)
```

# Table of sample sizes

```{r}
table_sample_sizes <- dat %>% 
  dplyr::summarise('studies' = n_distinct(study_ID),
                   'species' = n_distinct(Phylogeny),
                   'pfas type' = n_distinct(pfas_ID),
                   'Adult sample types' = n_distinct(sample_type_A),
                   'Progeny sample types' = n_distinct(sample_type_P),
                   'Breeding stage' = n_distinct(breeding_stage),
                   'Laying order' = n_distinct(laying_order),
                   'Same nest' = n_distinct(nest),
                   'effect sizes' = n_distinct(es_ID),
                   'effect sizes (eggs)' = n_distinct(es_ID[life_stage_P == "egg"]),
                   'studies (eggs)' = n_distinct(study_ID[life_stage_P == "egg"]),
                   'species (eggs)' = n_distinct(Phylogeny[life_stage_P == "egg"]),
                   'effect sizes (chicks)' = n_distinct(es_ID[life_stage_P == "chick"]),
                   'studies (chicks)' = n_distinct(study_ID[life_stage_P == "chick"]),
                   'species (chicks)' = n_distinct(Phylogeny[life_stage_P == "chick"])
                   )

table_sample_sizes<-t(table_sample_sizes)

colnames(table_sample_sizes)<-"n (sample size)"

kable(table_sample_sizes) %>% 
  kable_styling("striped", position="left")
```

# Summary of the dataset

```{r}
kable(summary(dat), "html") %>% 
  kable_styling("striped", position = "left") %>% 
    scroll_box(width = "100%", height = "500px")
```

# Meta-analysis {.tabset}

## Intercept meta-analytical model

Now that we calculated the effect size measures, we fit an intercept only, meta-analytic model

read <https://daniel1noble.github.io/orchaRd/> for more details on making pretty plots.

First of all, we need to account for correlated errors between cohorts (shared_controls)
```{r}
VCV_lnRR <- impute_covariance_matrix(dat$var_lnRR, cluster = dat$measurement_ID_A, r = 0.5)
```

```{r, eval=FALSE}
model1 <- rma.mv(yi = lnRR,
             V = VCV_lnRR,
             random = list(~1|es_ID,
                           ~1|study_ID),
             test = "t",
             data = dat)

save(model1, file = here("Rdata", "model1.RData"))
```

```{r}
load(here("RData", "model1.RData"))
summary(model1)
i2_ml(model1)
orchard_plot(model1, group = "study_ID", data = dat, xlab = "lnRR")
```

Use i2_sn function to obtain the total I\^2

```{r}
I2 <- orchaRd::i2_ml(model1)

orchaRd::orchard_plot(model1,
                      mod="1",
                      group = "study_ID",
                      data = dat,
                      xlab = "lnRR") + 
  annotate(geom="text",
           x= 0.80,
           y= 1, 
           label= paste0("italic(I)^{2} == ", round(I2[1],4)), # I want to use % but cannot do should come back
              color="black", parse = TRUE, size = 5) +
  scale_fill_manual(values="grey") +
  scale_colour_manual(values="grey")
```

From previous graphs we can see that overall estimate from a random-effects meta-analysis of 371 effect sizes is centered on zero, with a 95% CI that spans the line of no-effect. The prediction intervals clearly demonstrate the high level of heterogeneity (over 95%).

We can now draw a comparable caterpillar plot by using the function caterpillars.

```{r,eval=FALSE}
orchaRd::caterpillars(model1_res, mod="Int", xlab = "lnRR") 
```

## Determine the random effect structure

Now we want to assess what variable explain for variance in the model.

Test the random effects.

```{r, eval=FALSE}
model2 <- rma.mv(lnRR, VCV_lnRR, # Add `VCV_lnRR` to account for correlated errors between cohorts (shared_controls)
              random = list(~1|study_ID, # Identity of the study
                            ~1|Phylogeny, # Phylogenetic correlation between species
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              R = list(Phylogeny = cor_tree), # Assign the 'Phylogeny' argument to the phylogenetic variance-covariance matrix
              test = "t",
              data = dat,
              sparse = TRUE)

save(model2, file = here("Rdata", "model2.RData"))
```

```{r}
load(here("RData", "model2.RData"))
summary(model2)
i2_ml(model2)
```

Phylogeny does not explain for any variance. Thus, we remove it from the model.

## Meta-regression model
Now, we want to fit a meta-regression model to test whether there are moderators that explain for variance.

### Tissue progeny
Does "sample_type_P" impact our results on the mean and variance?

```{r, eval=FALSE}
model3 <- rma.mv(lnRR, VCV_lnRR,
              mod = ~ sample_type_P,# + sample_type_A,
              random = list(~1|study_ID, # Identity of the study
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              #R = list(Phylogeny = cor_tree), # Assign the 'Phylogeny' argument to the phylogenetic variance-covariance matrix
              test = "t",
              data = dat,
              sparse = TRUE)
save(model3, file = here("Rdata", "model3.RData"))
```

```{r}
load(here("RData", "model3.RData"))
summary(model3)
i2_ml(model3)
orchard_plot(model3, mod = "sample_type_P", group = "study_ID", data = dat, xlab = "lnRR")
```

### Tissue adult
Does "sample_type_A" impact our results on the mean and variance?

```{r, eval=FALSE}
model4 <- rma.mv(lnRR, VCV_lnRR,
              mod = ~ sample_type_A,
              random = list(~1|study_ID, # Identity of the study
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              #R = list(Phylogeny = cor_tree), # Assign the 'Phylogeny' argument to the phylogenetic variance-covariance matrix
              test = "t",
              data = dat,
              sparse = TRUE)
save(model4, file = here("Rdata", "model4.RData"))
```

```{r}
load(here("RData", "model4.RData"))
summary(model4)
i2_ml(model4)
orchard_plot(model4, mod = "sample_type_A", group = "study_ID", data = dat, xlab = "lnRR")
```

### Bird species 
Now, we want to fit a meta-regression model to test whether the moderator "species_ID" impacts our results on the mean and variance.

```{r, eval=FALSE}
model5 <- rma.mv(lnRR, VCV_lnRR,
              mod = ~ species_ID,
              random = list(~1|study_ID, # Identity of the study
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              test = "t",
              data = dat,
              sparse = TRUE)
save(model5, file = here("Rdata", "model5.RData"))

```

```{r}
load(here("RData", "model5.RData"))
summary(model5)
i2_ml(model5)
orchard_plot(model5, mod = "species_ID", group = "study_ID", data = dat, xlab = "lnRR")
```

### PFAS type - carboxylic acids

```{r, eval=FALSE}
model6 <- rma.mv(lnRR, VCV_lnRR,
              mod = ~ pfas_type_acronym,
              random = list(~1|study_ID, # Identity of the study
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              test = "t",
              data = dat,
              sparse = TRUE)
save(model6, file = here("Rdata", "model6.RData"))
```

```{r}
load(here("RData", "model6.RData"))
summary(model6)
i2_ml(model6)
orchard_plot(model6, mod = "species_ID", group = "study_ID", data = dat, xlab = "lnRR")
```

### PFAS type - sulfonic acids

```{r, eval=FALSE}
model7 <- rma.mv(lnRR, VCV_lnRR,
              mod = ~ pfas_type_acronym,
              random = list(~1|study_ID, # Identity of the study
                            ~1|species_ID,
                            ~1|pfas_ID, # Type of PFAS
                            ~1|es_ID # Effect size identity
              ),
              test = "t",
              data = dplyr::filter(dat,
                                   functional_group == "Sulfonate / sulfonic acid"),
              sparse = TRUE)
save(model7, file = here("Rdata", "model7.RData"))
```

```{r}
load(here("RData", "model7.RData"))
summary(model7)
i2_ml(model7)
orchard_plot(model7, mod = "species_ID", group = "study_ID", data = dat, xlab = "lnRR")
```







  
  
  
  









```{r, eval = FALSE}

mr2_egg <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ carbon_chain_length,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dplyr::filter(dat,
                                   functional_group == "Carboxylate / carboxylic acid" & life_stage_P == "egg"))

#summary(mr2_egg)

mr2_egg_res <- orchaRd::mod_results(mr2_egg, 
                             mod = "carbon_chain_length",
                             data = dplyr::filter(dat,
                                   functional_group == "Carboxylate / carboxylic acid" & life_stage_P == "egg"),
                             group = "study_ID",
                             )
mr2_egg_res

p2_egg <- orchaRd::orchard_plot(mr2_egg_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p2_egg

#orchaRd::caterpillars(mr2_egg_res, mod="Int", xlab = "lnRR")
```

###PFAS type - sulfonic acids

```{r, eval = FALSE}
mr3 <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ carbon_chain_length,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dplyr::filter(dat,
                                   functional_group == "Sulfonate / sulfonic acid"))

#summary(mr3)



mr3_res <- orchaRd::mod_results(mr3, 
                             mod = "carbon_chain_length",
                             data = dplyr::filter(dat,
                                   functional_group == "Sulfonate / sulfonic acid"),
                             group = "study_ID",
                             )
mr3_res



p3 <- orchaRd::orchard_plot(mr3_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p3

#orchaRd::caterpillars(mr3_res, mod="Int", xlab = "lnRR")

# Same model but only on eggs:

mr3_egg <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ carbon_chain_length,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dplyr::filter(dat,
                                   functional_group == "Sulfonate / sulfonic acid" & life_stage_P == "egg"))

#summary(mr3_egg)

mr3_egg_res <- orchaRd::mod_results(mr3_egg, 
                             mod = "carbon_chain_length",
                             data = dplyr::filter(dat,
                                   functional_group == "Sulfonate / sulfonic acid" & life_stage_P == "egg"),
                             group = "study_ID",
                             )
mr3_egg_res

p3_egg <- orchaRd::orchard_plot(mr3_egg_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p3_egg
```

###Seabirds VS non-seabirds

```{r, eval = FALSE}
mr4 <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ seabird,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dat)

#summary(mr4)

mr4_res <- orchaRd::mod_results(mr4, 
                             mod = "seabird",
                             data = dat,
                             group = "study_ID",
                             )
mr4_res

p4 <- orchaRd::orchard_plot(mr4_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p4

#orchaRd::caterpillars(mr4_res, mod="Int", xlab = "lnRR")

# Same model but only on eggs:

mr4_egg <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ seabird,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dplyr::filter(dat, life_stage_P == "egg"))

#summary(mr4_egg)

mr4_egg_res <- orchaRd::mod_results(mr4_egg, 
                             mod = "seabird",
                             data = dplyr::filter(dat, life_stage_P == "egg"),
                             group = "study_ID",
                             )
mr4_egg_res

p4_egg <- orchaRd::orchard_plot(mr4_egg_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p4_egg

#orchaRd::caterpillars(mr4_egg_res, mod="Int", xlab = "lnRR")
```

#Publication bias

```{r}
pb_model <- metafor::rma.mv(lnRR, 
              var_lnRR,
              mods = ~ year_publication,
              random = list(~1|es_ID,
                           ~1|study_ID),
              data = dat)
pb_bubble <- orchaRd::mod_results(pb_model,
                                  mod = "year_publication",
                                  group = "study_ID",
                                  data = dat)

orchaRd::bubble_plot(pb_bubble, 
                     data = dat,
                     group = "study_ID",
                     mod = "year_publication",
                     xlab = "Year", 
                     legend.pos = "top.left")
```

# Sensitivity analysis - Models

Here we run each model using data for sensitivity analysis (dat2 and dat3).

```{r}

```

# tries

Check extreme lnRR values

```{r}
extreme_values <- 
  dat %>%
  dplyr::filter(lnRR < -4 | lnRR > 4)
colnames(extreme_values)
extreme_values <- 
  extreme_values[, c(57, 58, 1 : 56)]
view(extreme_values)
extreme_values <-
  extreme_values %>%
  dplyr::select(-c(mean_A, mean_P, sd_A, sd_P, n_A, n_P))
```

```{r}
ggplot(data=dplyr::filter(dat, life_stage_P == "egg"),
       aes(x=sample_type_P, y=lnRR, color = sample_type_A)) +
  geom_point(size = 5) +
  theme_minimal()
```

```{r}
boxplot( lnRR ~ sample_type_P, data = dplyr::filter(dat, life_stage_P == "egg"),
        col = c("green","yellow","purple"))
```




```{r, eval = FALSE}
mr1_egg <- rma.mv(lnRR,
              var_lnRR,
              mods = ~ species_ID,
              random = list(~1|es_ID,
                           ~1|study_ID),
              test = "t",
              data = dplyr::filter(dat,
                                   life_stage_P == "egg"))
summary(mr1_egg)
```

Again, we can create a table of results and draw plots

```{r, eval = FALSE}

mr1_egg_res <- orchaRd::mod_results(mr1_egg, 
                             mod = "species_ID",
                             data = dplyr::filter(dat,
                                   life_stage_P == "egg"),
                             group = "study_ID",
                             )
mr1_egg_res

p1_egg <- orchaRd::orchard_plot(mr1_egg_res,
                            mod = "Int",
                            data = "Int",
                            group = "Int",
                            xlab = "lnRR",
                            alpha=0.4,
                            angle = 45)
p1_egg

#orchaRd::caterpillars(mr1_egg_res, mod="Int", xlab = "lnRR")
```