---
title: "Code"
author: "LR"
date: '2022-06-28'
output: html_document
---

# The code of the following chunck merges the various datasheets of my relational database to create a final whole datasheet.

```{r set up and loading files}
#set up and packages loading
library(readr)
library(dplyr)
library(here)
library(ggplot2)
#dataframes loading
study_info <- read_csv(here("Data", "study_info.csv"))
species_info <- read_csv(here("Data", "species_info.csv"))
pfas_info <- read_csv(here("Data", "pfas_info.csv"))
cohort_info <- read_csv(here("Data", "cohort_info.csv"))
sample_info <- read_csv(here("Data", "sample_info.csv"))
measurement <- read_csv(here("Data", "measurement.csv"))
#merging measurement dataframe and sample dataframe
ms <- inner_join(measurement, sample_info, by = "sample_ID")
msc <- inner_join(ms, cohort_info, by = "cohort_ID")
mscs <- inner_join(msc, study_info, by = "study_ID")
mscsp <- inner_join(mscs, pfas_info, by = "pfas_ID")
mscsps <- inner_join(mscsp, species_info, by = "species_ID")
data_combined <- mscsps %>% 
  select(-c(Timestamp.x, Timestamp.y, Timestamp.x.x, Timestamp.y.y, Timestamp.x.x.x, Timestamp.y.y.y))
View(data_combined)
```
# Clean the dataframe removing comment and statement columns
```{r cleaning}
data_combined <- data_combined %>% 
  select(- ends_with("comment")) %>% 
  select(- ends_with("statement"))
```

## To calculate the overall [PFAS]adult / [PFAS]progeny ratio I need to devide the mean of egg for the mean of adult (group_info).
```{r}
#progeny <- filter(data_combined, group_info == "egg" | group_info == "chick")
#adult <- filter(data_combined, group_info == "adult")
#ggplot(data_combined) %>% 
  #geom_point(aes(x=group_by() mean_arithmetic, y=adult$mean_arithmetic))
```

## Summary of scientific and common names of the species
```{r}
data_combined %>% group_by(species_scientific_name, species_common_name) %>% summarise()
```


Create a new dataframe without not necessary data for MTR calculation:

```{r}
data_MTR <- data_combined %>%
  select(Measurement_ID, Sample_ID, pfas_ID, Mean, SE, SD, N, group_info, Cohort_ID, Sample_type, Species_ID, Study_ID)
```
Remove S_002 because it doesn't have clear data yet

```{r}
data_MTR <- data_MTR %>%
  filter(Study_ID != "S_002")
```
Overall meta-analytic mean calculation:
```{r}

```

