---
title: "Creating a dm object"
author: "LR"
date: "2022-07-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

We use dm to create a relational data model from our local data frames.
The six tables that we are working with contain data extracted from the
included studies: **study_info**, **species_info**, **pfas_info**,
**cohort_info**, **sample_info**, **measurements**.

```{r loading packages}
library(readr)
library(here)
library(dm)
library(dplyr)
```

```{r loading data, include=FALSE}
st <- study_info <- read_csv(here("Data", "study_info.csv"))
sp <- species_info <- read_csv(here("Data", "species_info.csv"))
pfas <- pfas_info <- read_csv(here("Data", "pfas_info.csv"))
co <- cohort_info <- read_csv(here("Data", "cohort_info.csv"))
sa <- sample_info <- read_csv(here("Data", "sample_info.csv"))
me <- measurements <- read_csv(here("Data", "measurement.csv"))
```

```{r cleaning - removing unnecessary columns (i.e., "Timestamp" and comment columns)}
st <- st %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
sp <- sp %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
pfas <- pfas %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
co <- co %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
sa <- sa %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
me <- me %>% select(-("Timestamp")) %>% select(-(ends_with("comment")))
```

```{r adding tables, include=FALSE}
data_dm_no_keys <- dm(st, sp, pfas, co, sa, me)
data_dm_no_keys
data_dm_no_keys$st
data_dm_no_keys[c("st", "co")]
```

```{r defining primary keys}
data_dm_only_pks <- data_dm_no_keys %>% 
  dm_add_pk(table = st, columns = study_ID) %>% 
  dm_add_pk(sp, species_ID) %>% 
  dm_add_pk(pfas, pfas_ID) %>% 
  dm_add_pk(co, cohort_ID) %>% 
  dm_add_pk(sa, sample_ID) %>% 
  dm_add_pk(me, measurement_ID)
data_dm_only_pks
```

```{r defining foreign keys}
data_dm_all_keys <- 
  data_dm_only_pks %>% 
  dm_add_fk(table = co, columns = study_ID, ref_table = st) %>% 
  dm_add_fk(table = co, columns = species_ID, ref_table = sp) %>%
  dm_add_fk(table = sa, columns = cohort_ID, ref_table = co) %>%
  dm_add_fk(table = me, columns = sample_ID, ref_table = sa) %>%
  dm_add_fk(table = me, columns = pfas_ID, ref_table = pfas)
data_dm_all_keys
```

```{r visualization}
data_dm_all_keys %>% 
  dm_draw()
```

```{r integrity check}
data_dm_all_keys %>% 
  dm_examine_constraints()
```

```{r joining example - moving data between dm data tables. Add a column.}
# Add a column with study year of publication from the study_info table to the cohort_info table
dm_co_plus_year <- 
  data_dm_all_keys %>% 
  dm_select(st, study_ID | year_publication) %>% 
  dm_flatten_to_tbl(co, st, .join = left_join)
dm_co_plus_year
```

```{r merging all tables into one big table}
dm_joined <- 
  data_dm_all_keys %>% 
  dm_flatten_to_tbl(.start = me, .recursive = TRUE)
dm_joined
```

Zooming and manipulating tables

```{r adding a new column to an existing table - n > or < 10}
me_zoomed <- 
  data_dm_all_keys %>% 
  dm_zoom_to(me)
me_plus_n_info <- 
  me_zoomed %>% 
  mutate(n_morethan_ten = if_else(n > 10, "true", "false")) %>%
  select(measurement_ID:n, n_morethan_ten, everything()) # in order to see our changes in the output we use `select()` for reordering the columns
me_plus_n_info$n_morethan_ten
# To update the original `dm` with a new `me` table we use `dm_update_zoomed()`
updated_data_dm <- 
  me_plus_n_info %>% 
  dm_update_zoomed()
updated_data_dm # from 56 to 57 columns
dm_draw(updated_data_dm) # The schematic view of the data model remains unchanged
```

Adding a summary of a table as a new table to a dm. Adding to dm a table
with the mean pfas concentrations of each pfas type.

```{r adding a summary table to dm}
dm_with_summary <- 
  updated_data_dm %>% 
  dm_zoom_to(me) %>% 
  group_by(pfas_ID) %>% 
  summarise(mean = mean(mean_arithmetic, na.rm = TRUE)) %>% 
  dm_insert_zoomed("pfas_mean")
dm_draw(dm_with_summary)
dm_with_summary$pfas_mean
```

```{r look at a single dm table}
tbl(data_dm_all_keys, "sp")
```


Filtering data - we want to filter our data with only the species Larus argentatus.
```{r filtering data}
filtered_dm <- 
  data_dm_all_keys %>% 
  dm_filter(sp = (species_scientific_name == "Larus argentatus"))
dm_nrow(filtered_dm) # sp 1 row, me 124 rows
```









