---
title: "deduplication"
author: "LR"
date: '2022-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Deduplication workflow. The file named "articles_pilot_2.csv" contains the output of the searching process of databases. We merged the articles found searching databases using Rayyan.

### Setup and packages loading
```{r, include=FALSE}
library(tidyverse)
library(here)
library(synthesisr)
library(tidystringdist)
library(bibliometrix)
```
### Loading file
```{r, include=FALSE}
dat <- read_csv(here("Data", "articles_pilot_2.csv"))
dim(dat) #[1] 1244   19
```
### Tidy up and simplify titles removing all punctuation and extra white spaces
```{r, include=FALSE}
dat$title2 <- str_replace_all(dat$title,"[:punct:]","") %>% str_replace_all(.,"[ ]+", " ") %>% tolower()
```
### Remove exact title matches
```{r, include=FALSE}
dat2 <- distinct(dat, title2, .keep_all = TRUE) #select records with unique titles (removes exact duplicates)
dim(dat2) #[1] 860  20
```
### Removing partial matches in titles
```{r, include=FALSE}
duplicates_string <- find_duplicates(dat2$title2, method = "string_osa", to_lower = TRUE, rm_punctuation = TRUE, threshold = 7)
```
### Manually review those titles to confirm they are duplicates
```{r, include=FALSE}
manual_checks <- review_duplicates(dat2$title, duplicates_string)
view(manual_checks)
dat3 <- extract_unique_references(dat2, duplicates_string)
dim(dat3) #[1] 827  21
names(dat3)
```
### Drop columns "title2" and "n_duplicates"
```{r, include=FALSE}
dat4 <- select(dat3, -c(title2,n_duplicates))
dim(dat4) #[1] 827  19
```
### # Save as a .csv file and .bib file
```{r, include=FALSE}
write_csv(dat4, "abstracts_for_screening_deduplicated.csv")
write_refs(dat4, format = "bib", file = "abstracts_for_screening_deduplicated.bib")
```
